### 2018-08-07 article.tex
minor edits

\section{Mathematical Preliminaries}
\label{sec:prelim}

@@

PROPs have been used to formalise many algebraic structures
\cite{Lack:2004sf}, however our interest here is to formalise a
general notion of \emph{circuit} as a free PROP over a certain set of
gates.

@@

\begin{example} \label{ex:quantum-circs}
  The main example of interest is \Circ, the PROP of quantum circuits.
  Different authors, and different hardware manufacturers, prefer
  different primitive gate sets but the following is typical:
  \[
  \begin{array}{cccccccccc}
    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
    \vstrut{1.2em}
    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
    \\
    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
    \vstrut{1.2em}
    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
  \end{array}
  \]
  We omit the usual matrix interpretations but these can be found in
  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
  this example we actually have an uncountable family of generating
  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
  Boolean circuits all quantum gates have the same number of inputs and
  outputs.  In fact, \Circ is a \dag-PROP, but this will not concern
  us here.
\end{example}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        MONOIDAL 2-CATEGORY DEFINITION        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{definition}\label{def:monoidal-two-cat}
  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a strict
    2-category with a 2-bifunctor $\otimes: \mathcal{C} \times \mathcal{C} \to
    \mathcal{C}$.
  \REM{This definition doesn't quite work without the additoin type
    information on the \emph{2-cells}}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        INTERCHANGE LAW REMARK        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}[Interchange law]
  Since $\circ$ and $\otimes$ are bifunctorial, we have the following
    interchange equations for all \emph{0-cells} $A, B \in \mathcal{C}$ and all
    \emph{2-cells} $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
  \begin{align}
    (\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet
      \gamma) \circ (\beta \bullet \delta) \\
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
      \gamma) \otimes (\beta \bullet \delta)
  \end{align}
  Since $\otimes$ is a 2-functor, we have for all \emph{0-cells} $A, B \in
    \mathcal{C}$ and all \emph{2-cells} $\alpha, \beta, \gamma, \delta \in
    \mathcal{C}(A, B)$
  \[
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) = (\alpha \bullet
      \gamma) \bullet (\beta \bullet \delta)
  \]
\end{remark}

>>>>

\section{Circuits and 2-PROPS}
\label{sec:prelim}

@@

PROPs have been used to formalise many algebraic structures
\cite{Lack:2004sf}, however our interest here is to formalise a
general notion of \emph{circuit} made by wiring together certain
atomic operations.  With this in mind, a PROP of circuits is
henceforth a PROP freely generated from some monoidal signature; the
members of the signature are called \emph{gates}, and a circuit is
a morphism in such a PROP.

@@

\begin{example} \label{ex:quantum-circs}
  The main example of interest is \Circ, the PROP of quantum circuits.
  Different authors, and different hardware manufacturers, prefer
  different primitive gate sets but the following is typical:
  \[
  \begin{array}{cccccccccc}
    \circZ{{}} & \qquad  & \circX{{}} & \qquad & \circH \\ 
    \vstrut{1.2em}
    Z : 1 \to 1 && X : 1 \to 1 && H : 1 \to 1\\
    \\
    \circbox{R_Z(\theta)} & \qquad & \circCPhase{\theta} & \qquad & \circCX  \\ 
    \vstrut{1.2em}
    R_Z(\theta) : 1 \to 1 &&  \CPhase\theta :2 \to 2 && \CX : 2 \to 2\\
  \end{array}
  \]
  We omit the usual matrix interpretations but these can be found in
  the standard textbook \cite{NieChu:QuantComp:2000}.  Note that in
  this example we actually have an uncountable family of generating
  gates, one for each value $\theta \in [0,2\pi)$.  Unlike the case of
  Boolean circuits all quantum gates have the same number of inputs and
  outputs.  In fact, \Circ is a \dag-PROP and all its but this will not concern
  us here.
\end{example}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        MONOIDAL 2-CATEGORY DEFINITION        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{definition}\label{def:monoidal-two-cat}
  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a
  strict 2-category with a 2-bifunctor $\otimes: \mathcal{C} \times
  \mathcal{C} \to \mathcal{C}$.
  \TODO{Denis: fix this definition}
  \REM{What about the unit object?}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        INTERCHANGE LAW REMARK        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}[Interchange law]
\TODO{DENIS : please fix this statement.  THe second and third clauses
  are the same.!}
  Since $\circ$ and $\otimes$ are bifunctorial, we have the following
    interchange equations for all \emph{0-cells} $A, B \in \mathcal{C}$ and all
    \emph{2-cells} $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
  \begin{align}
    (\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet
      \gamma) \circ (\beta \bullet \delta) \\
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
      \gamma) \otimes (\beta \bullet \delta)
  \end{align}
  Since $\otimes$ is a 2-functor, we have for all \emph{0-cells} $A, B \in
    \mathcal{C}$ and all \emph{2-cells} $\alpha, \beta, \gamma, \delta \in
    \mathcal{C}(A, B)$
  \[
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) = (\alpha \bullet
      \gamma) \bullet (\beta \bullet \delta)
  \]
\end{remark}

Putting the preceding definitions together, a 2-PROP is then a strict
monoidal 2-category whose underlying monoidal 1-category is a PROP, or
equivalently, a strict monoidal 2-category whose 0-cells are generated
by a single object under $\otimes$.
